### 2016-06-21 paper/icfp16-cameraready/introduction.tex

\subparagraph{Witness Utility}
%
Even if we can find small witnesses for the majority of type errors, it
may be that the witnesses do not actually help developers
\emph{understand} the errors.
%
In other words, perhaps the static error message is sufficient to
diagnose and fix the error, or perhaps the witness simply does not add
enough information to make a difference.
%
Thus, our final contribution is a user study that compares the utility
of our witnesses with that of the error messages provided by the \ocaml
compiler~(\S~\ref{sec:user-study}).
%
We show that students given one of our witnesses are consistently more
likely to correctly explain and fix a type error than those given
the standard error message produced by the \ocaml compiler.

\smallskip
Our results show that in the vast majority of
cases, (novices') ill-typed programs \emph{do} go wrong,
and that the witnesses to these errors can be helpful in
understanding the source of the error.
%
This, in turn, opens the door to a novel dynamic way to
explain, understand, and appreciate the benefits of static typing.

>>>>

The ultimate purpose of an error report is to help the programmer 
\emph{comprehend} and \emph{fix} problematic code.
%
Thus, our final contribution is a user study that compares \toolname's 
dynamic witnesses against \ocaml's type errors along the dimension of
comprehensibility~(\S~\ref{sec:user-study}).
%
Our study finds that students given one of our witnesses are 
consistently more likely to correctly explain and fix a type
error than those given the standard error message produced by 
the \ocaml compiler.

%
% \subparagraph{Witness Utility}
%
% Even if we can find small witnesses for the majority of type errors, it
% may be that the witnesses do not actually help developers
% \emph{understand} the errors.
%
% In other words, perhaps the static error message is sufficient to
% diagnose and fix the error, or perhaps the witness simply does not add
% enough information to make a difference.
%
%
% Thus, our final contribution is a user study that compares the utility
% of our witnesses with that of the error messages provided by the \ocaml
% compiler~(\S~\ref{sec:user-study}).
%

\smallskip
All together, our results show that in the vast majority of cases, (novices') ill-typed 
programs \emph{do} go wrong, and that the witnesses to these errors can be 
helpful in understanding the source of the error. This, in turn, opens the 
door to a novel dynamic way to explain, understand, and appreciate the 
benefits of static typing.
